---
title: "COVID-19 in India"

output: 
  flexdashboard::flex_dashboard:
    vertical_layout:  fill
    social: ["twitter", "facebook", "linkedin"]
    theme: readable
    
runtime: shiny
---

```{r setup, include=FALSE}

library(flexdashboard)
library(rjson)
library(DT)
library(dplyr)
library(lubridate)
library(deSolve)
library(geojsonio)
library(plotly)
library(ggplot2)
library(ggiraph)
library(colormap) 
library(hrbrthemes) 

# SIR model function
SIR <- function(time, state, parameters) {
  par <- as.list(c(state, parameters))
  with(par, {
    dS <- -beta * I * S / N
    dI <- beta * I * S / N - gamma * I
    dR <- gamma * I
    list(c(dS, dI, dR))
  })
}

result = fromJSON(file = "https://api.covid19india.org/data.json")

DeltaConfirmed = result$statewise[[1]]$deltaconfirmed
DeltaDeaths = result$statewise[[1]]$deltadeaths
DeltaRecovered = result$statewise[[1]]$deltarecovered

totalConfirmed = as.numeric(result$statewise[[1]]$confirmed)
totalActive = as.numeric(result$statewise[[1]]$active)
totalRecovered = as.numeric(result$statewise[[1]]$recovered)
totalDeath = as.numeric(result$statewise[[1]]$deaths)
totalSamplesTested = as.numeric(result$tested[[length(result$tested)]]$totalsamplestested)

india = data.frame(matrix(ncol = 7, nrow = 0))
colnames(india) = c("State/UT", "Confirmed", "Active", "Recovered", "Death", "Recovery Rate", "Death Rate")

for (i in 2:38)
  if(as.numeric(result$statewise[[i]]$confirmed)>0)
    india[nrow(india)+1, ] = list(result$statewise[[i]]$state, 
                                as.numeric(result$statewise[[i]]$confirmed), 
                                as.numeric(result$statewise[[i]]$active), 
                                as.numeric(result$statewise[[i]]$recovered), 
                                as.numeric(result$statewise[[i]]$deaths),
                                as.numeric(result$statewise[[i]]$recovered) / as.numeric(result$statewise[[i]]$confirmed),
                                as.numeric(result$statewise[[i]]$deaths) / as.numeric(result$statewise[[i]]$confirmed))

india = india[order(-india[, 2]), ]

copy = data.frame(india)
names(copy)[names(copy) == "State.UT"] <- "id"

shape = geojsonio::geojson_read("india-polygon.geojson", what = "sp")
states <- fortify(shape, region = "ST_NM")

merged_shape = merge(states, copy, by = "id", all.x = TRUE)
final_plot = merged_shape[order(merged_shape$order), ]
final_plot[is.na(final_plot)] = 0


# daily = data.frame(matrix(ncol = 7, nrow = 0))
# colnames(daily) = c("Date", "DailyConfirmed", "DailyRecovered", "DailyDeath", "TotalConfirmed", "TotalRecovered", "TotalDeath")
# 
# for(i in 1:length(result$cases_time_series))
#   daily[nrow(daily)+1, ] = list(format(as.Date(paste(trimws(result$cases_time_series[[i]]$date), "2020"), format = "%d %B %y"), "%d %b %Y"),
#                                 as.numeric(result$cases_time_series[[i]]$dailyconfirmed),
#                                 as.numeric(result$cases_time_series[[i]]$dailyrecovered),
#                                 as.numeric(result$cases_time_series[[i]]$dailydeceased),
#                                 as.numeric(result$cases_time_series[[i]]$totalconfirmed),
#                                 as.numeric(result$cases_time_series[[i]]$totalrecovered),
#                                 as.numeric(result$cases_time_series[[i]]$totaldeceased))

daily = data.frame(matrix(ncol = 9, nrow = 0))
colnames(daily) = c("Date", "DailyConfirmed", "DailyActive", "DailyRecovered", "DailyDeath", "TotalConfirmed", "TotalActive", "TotalRecovered", "TotalDeath")

for(i in 1:length(result$cases_time_series))
  daily[nrow(daily)+1, ] = list(format(as.Date(paste(trimws(result$cases_time_series[[i]]$date), "2020"), format = "%d %B %y"), "%d %b %Y"),
                                as.numeric(result$cases_time_series[[i]]$dailyconfirmed),
                                as.numeric(result$cases_time_series[[i]]$dailyconfirmed) - as.numeric(result$cases_time_series[[i]]$dailyrecovered) - as.numeric(result$cases_time_series[[i]]$dailydeceased),
                                as.numeric(result$cases_time_series[[i]]$dailyrecovered),
                                as.numeric(result$cases_time_series[[i]]$dailydeceased),
                                as.numeric(result$cases_time_series[[i]]$totalconfirmed),
                                as.numeric(result$cases_time_series[[i]]$totalconfirmed) - as.numeric(result$cases_time_series[[i]]$totalrecovered) - as.numeric(result$cases_time_series[[i]]$totaldeceased),
                                as.numeric(result$cases_time_series[[i]]$totalrecovered),
                                as.numeric(result$cases_time_series[[i]]$totaldeceased))

```

Summary {data-orientation=rows data-icon="fa-list-alt"}
===========

Row{}
----------------------

### Total Samples tested

```{r}

valueBox(value = paste(format(totalSamplesTested, big.mark = ","), "", sep = ""), icon = "fa-flask")

```

### Total Confirmed cases

```{r}

if(DeltaConfirmed>=0) {
  valueBox(value = paste(format(totalConfirmed, big.mark = ","), " (", "\u25B2", DeltaConfirmed, ")", sep = ""), icon = "fa-check-circle")
} else {
  valueBox(value = paste(format(totalConfirmed, big.mark = ","), " (", "\u25BC", DeltaConfirmed, ")", sep = ""), icon = "fa-check-circle")
}

```

### Total Active cases

```{r}

valueBox(value = paste(format(totalActive, big.mark = ","), "", sep = ""), icon = "fa-toggle-on")

```

### Total Recovered

```{r}

if(DeltaRecovered>=0) {
  valueBox(value = paste(format(totalRecovered, big.mark = ","), " (", "\u25B2", DeltaRecovered, ")", sep = ""), icon = "fa-heartbeat")
} else {
  valueBox(value = paste(format(totalRecovered, big.mark = ","), " (", "\u25BC", DeltaRecovered, ")", sep = ""), icon = "fa-heartbeat")
}
  

```

### Total Deaths

```{r}

if(DeltaDeaths>=0) {
  valueBox(value = paste(format(totalDeath, big.mark = ","), " (", "\u25B2", DeltaDeaths, ")", sep = ""), icon = "fa-bed")
} else {
  valueBox(value = paste(format(totalDeath, big.mark = ","), " (", "\u25BC", DeltaDeaths, ")", sep = ""), icon = "fa-bed")
}
  

```


Row {}
-----------------------------------------------------------------------

### State wise details

```{r}

india %>%
  DT::datatable(rownames = FALSE,
                colnames = colnames(india),
                options = list(pageLength = nrow(india), dom = 'tip')) %>%
  DT::formatPercentage("Recovery Rate", 2) %>%
  DT::formatPercentage("Death Rate", 2) 

```


Visualization {data-icon="fa-bar-chart"}
==========

Column
------------------------

### State wise distribution of cases

```{r}

india$`State/UT` <- factor(india$`State/UT`, levels = unique(india$`State/UT`)[order(india$Confirmed, decreasing = TRUE)])

plotly::plot_ly(data = india,
                x = india$`State/UT`,
                y = india$Active, 
                type = "bar", 
                name = "Active",
                marker = list(color = "#1f77b4")) %>%
  
  plotly::add_trace(y = india$Recovered, 
                    name = "Recovered",
                    marker = list(color = "forestgreen")) %>%
    
  plotly::add_trace(y = india$Death, 
                    name = "Death",
                    marker = list(color = "red")) %>%
  
  plotly::layout(barmode = 'stack',
                 yaxis = list(title = "Total Cases"), 
                 hovermode = "compare",
                  margin =  list(
                   b = 10,
                   t = 10,
                   pad = 2
                 ),
                 legend = list(x = 0.1, y = 1.4))

```

Column {}
-----------------------------------------------------------------------

### India map

```{r}

renderggiraph({
  
  g <- ggplot(final_plot) +
  geom_polygon_interactive(
    color='black', size = 0.3, 
    aes(x = long, y = lat, group = group, fill = Confirmed,
        tooltip = sprintf("State/UT: %s <br> Confirmed: %s <br> Active: %s <br> Recovered: %s <br> Death(s): %s", id, Confirmed, Active, Recovered, Death))) +
  guides(fill=guide_legend("No. of cases")) +
  hrbrthemes::theme_ipsum() +
  colormap::scale_fill_colormap(
    colormap=colormap::colormaps$velocity_blue, reverse = T) +
  labs(title='Total cases in India', subtitle = "Hover over a state for more details")

  girafe_options(ggiraph(code = print(g), width_svg = 12, height_svg = 12))
})

```

Trends {data-orientation=rows data-icon="fa-chart-line"}
==========

Row {.tabset}
-------------------

### Cumulative

```{r}

daily$Date = dmy(daily$Date)

# plotly::plot_ly(data = daily) %>%
#   
#   plotly::add_trace(x = daily$Date , 
#                     y = daily$TotalConfirmed,
#                     type = "scatter",
#                     mode = "lines+markers",
#                     name = "Confirmed",
#                     line = list(color = "#1f77b4"),
#                     marker = list(color = "#1f77b4")) %>%
# 
#   plotly::add_trace(x = daily$Date , 
#                     y = daily$TotalRecovered,
#                     type = "scatter",
#                     mode = "lines+markers",
#                     name = "Recovered",
#                     line = list(color = "forestgreen"),
#                     marker = list(color = "forestgreen")) %>%
#   
#   plotly::add_trace(x = daily$Date , 
#                     y = daily$TotalDeath,
#                     type = "scatter",
#                     mode = "lines+markers",
#                     name = "Death",
#                     line = list(color = "red"),
#                     marker = list(color = "red")) %>%
#   
#   plotly::layout(title = "",
#                  yaxis = list(title = "Cumulative cases"),
#                  xaxis = list(title = "Date"),
#                  legend = list(x = 0.1, y = 0.9),
#                  hovermode = "compare")

plotly::plot_ly(data = daily,
                x = daily$Date,
                y = daily$TotalActive, 
                name = 'Active', 
                fillcolor = "#1f77b4",
                type = 'scatter',
                mode = 'none', 
                stackgroup = 'one') %>%
  
  plotly::add_trace(y = daily$TotalRecovered,
                    name = "Recovered",
                    fillcolor = "forestgreen") %>%
  
  plotly::add_trace(y = daily$TotalDeath,
                    name = "Death",
                    fillcolor = "red") %>%
  
  plotly::layout(title = "",
                 yaxis = list(title = "Cumulative Number of Cases"),
                 xaxis = list(title = "Date"),
                 legend = list(x = 0.1, y = 0.9),
                 hovermode = "compare")

```

### Daily

```{r}

#daily$Date = dmy(daily$Date)

plotly::plot_ly(data = daily) %>%
  
  plotly::add_trace(x = daily$Date , 
                    y = daily$DailyActive,
                    type = "scatter",
                    mode = "lines+markers",
                    name = "Active",
                    line = list(color = "#1f77b4"),
                    marker = list(color = "#1f77b4")) %>%

  plotly::add_trace(x = daily$Date , 
                    y = daily$DailyRecovered,
                    type = "scatter",
                    mode = "lines+markers",
                    name = "Recovered",
                    line = list(color = "forestgreen"),
                    marker = list(color = "forestgreen")) %>%
  
  plotly::add_trace(x = daily$Date , 
                    y = daily$DailyDeath,
                    type = "scatter",
                    mode = "lines+markers",
                    name = "Death",
                    line = list(color = "red"),
                    marker = list(color = "red")) %>%
  
  plotly::layout(title = "",
                 yaxis = list(title = "Daily cases"),
                 xaxis = list(title = "Date"),
                 legend = list(x = 0.1, y = 0.9),
                 hovermode = "compare")

```

SIR Model {.storyboard data-orientation=rows data-icon="fa-area-chart"}
======================

### COVID-19 Predicted vs Observed cases, India

```{r}

Infected = daily$TotalActive
Infected=Infected[c(33:length(Infected))] # From 2nd March 2020 till previous day
Day <- 1:(length(Infected))

N = 1380004385 # Population of India
init = c(S = N - Infected[1], I = Infected[1], R = 0)

RSS <- function(parameters) {
  names(parameters) <- c("beta", "gamma")
  out <- ode(y = init, times = Day, func = SIR, parms = parameters)
  fit <- out[, 3]
  sum((Infected - fit)^2)
}

Opt <- optim(c(0.5, 0.5), RSS, method = "L-BFGS-B", lower = c(0, 0), upper = c(1, 1), control = list(parscale = c(10^-4,10^-4)))
Opt_par <- setNames(Opt$par, c("beta", "gamma"))

sir_start_date <- "2020-03-02"

t <- 1:as.integer(ymd(Sys.Date()) - ymd(sir_start_date))

fitted_cumulative_incidence <- data.frame(ode(y = init, times = t, func = SIR, parms = Opt_par))

fitted_cumulative_incidence <- fitted_cumulative_incidence %>%
  mutate(
    Date = ymd(sir_start_date) + days(t - 1),
    Country = "India",
    cumulative_incident_cases = Infected
  )

fitted_cumulative_incidence %>%
  ggplot(aes(x = Date)) +
  geom_line(aes(y = I), colour = "red") +
  geom_point(aes(y = cumulative_incident_cases), colour = "blue") +
  labs(
    y = "Cumulative cases"
  ) +
  theme_minimal()

```

***

**SIR Model**

The basic idea behind the **SIR (Susceptible - Infectious - Recovered) model** of communicable disease outbreaks is that there are three groups (also called compartments) of individuals:

- **S**: those who are healthy but **susceptible** to the disease (i.e., at risk of being contaminated). At the start of the pandemic, S is the entire population since no one is immune to the virus.

- **I**: the **infectious** (and thus, infected) people

- **R**: individuals who were contaminated but who have either **recovered** or died. They are not infectious anymore.

**Legend:**

- Red - Predicted cases from SIR model

- Blue - Actually Observed cases

**Note:**

For the model, cases from 2nd March 2020 are only taken. 


### COVID-19 Predicted vs Observed cases, India (Log-linear plot)

```{r}

fitted_cumulative_incidence %>%
  ggplot(aes(x = Date)) +
  geom_line(aes(y = I), colour = "red") +
  geom_point(aes(y = cumulative_incident_cases), colour = "blue") +
  labs(y = "Cumulative cases") +
  theme_minimal() +
  scale_y_log10(labels = scales::comma)

```

***

**Legend:**

- Red - Predicted cases from SIR model

- Blue - Actually Observed cases

**Note:**

For the model, cases from 2nd March 2020 are only taken. 


### COVID-19 Predicted cases, India

```{r}

# Basic Reproduction number
R0 = as.numeric(Opt_par[1] / Opt_par[2])

t <- 1:150 # 150 days

# Get the fitted values from SIR model
fitted_cumulative_incidence <- data.frame(ode(
  y = init, times = t,
  func = SIR, parms = Opt_par
))

# With date
fitted_cumulative_incidence <- fitted_cumulative_incidence %>%
  mutate(
    Date = ymd(sir_start_date) + days(t - 1),
    Country = "India",
    cumulative_incident_cases = c(Infected, rep(NA, length(t) - length(Infected)))
  )

# Normal plot of data
fitted_cumulative_incidence %>%
  ggplot(aes(x = Date)) +
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = S), colour = "black") +
  geom_line(aes(y = R), colour = "green") +
  geom_point(aes(y = cumulative_incident_cases), colour = "blue") +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Cumulative cases", x = "Month", title = paste("R0 in India (according to the model) - ", sprintf(R0, fmt = '%.4f'))) +
  scale_colour_manual(name = "", values = c(red = "red", black = "black", green = "green", blue = "blue"), 
  labels = c("Susceptible", "Recovered", "Observed cases", "Infectious")
  ) +
  theme_minimal()

```

***

Known as the **Basic Reproduction number**, **R0** (pronounced “R-naught”) is the rate at which a virus is transmitted. It indicates the **average number of people who will contract the virus from a person who has already been infected**, in a population that does not have immunity for the said disease. 

**The higher the R0, the more contagious the infection.**

**Please note that this is a worst case prediction and should be taken with a lot of caution. Not surprisingly, it can be wrong. On one hand, this is based on rather unrealistic assumptions (for example, no public health interventions, fixed Reproduction number, etc.). On the other hand, we still have to be careful and strictly follow public health interventions because previous pandemics such as H1N1 and Spanish flu have shown that incredibly high numbers are not impossible! **

**Legend:**

- Black - Susceptible

- Blue - Observed cases

- Red - Infectious

- Green - Recovered

**Note:**

For the model, cases from 2nd March 2020 are only taken. 

No. of days taken for prediction - 150


### COVID-19 Predicted cases, India (Log-linear plot)

```{r}

# Logarithm plot
fitted_cumulative_incidence %>%
  ggplot(aes(x = Date)) +
  geom_line(aes(y = I, colour = "red")) +
  geom_line(aes(y = S, colour = "black")) +
  geom_line(aes(y = R, colour = "green")) +
  geom_point(aes(y = cumulative_incident_cases, colour = "blue")) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "Cumulative cases", x = "Month", title = paste("R0 in India (according to the model) - ", sprintf(R0, fmt = '%.4f'))) +
  scale_colour_manual(name = "", values = c(red = "red", black = "black", green = "green", blue = "blue"),
    labels = c("Susceptible", "Observed cases", "Recovered", "Infectious")
  ) +
  theme_minimal()

```

***

**Please note that this is a worst case prediction and should be taken with a lot of caution. Not surprisingly, it can be wrong. On one hand, this is based on rather unrealistic assumptions (for example, no public health interventions, fixed Reproduction number, etc.). On the other hand, we still have to be careful and strictly follow public health interventions because previous pandemics such as H1N1 and Spanish flu have shown that incredibly high numbers are not impossible! **

**Note:**

For the model, cases from 2nd March 2020 are only taken. 


About and FAQ {data-icon="fa-question-circle"}
==========

### About

**COVID-19 in India**

This dashboard/tracker gives an overview of the spread of the Novel Coronavirus (COVID-19) pandemic across India, state-wise along with **SIR (Susceptible - Infectious - Recovered) model** for a period of 150 days starting from March 2. 

This application has been made entirely using Rmarkdown framework.

**R packages used: **

* Dashboard - [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) package.

* Table - [DT](https://rstudio.github.io/DT/) and [dplyr](https://dplyr.tidyverse.org/) packages.

* API Data - [rjson](https://www.rdocumentation.org/packages/rjson/versions/0.2.20) package.

* Modeling - [deSolve](http://desolve.r-forge.r-project.org/) and [lubridate](https://lubridate.tidyverse.org/) packages. 

* Visualizations - [plotly](https://plot.ly/r/), [ggplot2](https://ggplot2.tidyverse.org/), [geojsonio](https://ropensci.org/tutorials/geojsonio_tutorial/), [ggiraph](https://davidgohel.github.io/ggiraph/), [colormap](https://bhaskarvk.github.io/colormap/) and [hrbrthemes](https://hrbrmstr.github.io/hrbrthemes/) packages. 

### FAQ

**Are you official?**

No.

**Who are you?**

I am an enthusiastic developer and a data analyst who wanted to do contribute to the fight against COVID-19.

**Is the project open sourced?**

Yes! The project is completely open sourced. You can find the link to the GitHub repo [here](https://github.com/srinathkr07/COVID-19-in-India/).

**From where do you get the data?**

Data collected by a group of volunteers is validated and published into a Google sheet and an API. Thanks for the API, [COVID-19 India API](https://api.covid19india.org/)! 

**Is the SIR model accurate?**

I ain't a specialist in Epidemiology so there are bound to be differences between actual and predicted cases. These predictions are only meant to be taken as a worst case scenario and with a lot of caution. 

**Created by [Srinath K R](https://www.linkedin.com/in/srinath-k-r-026147173/).**  
